Leader functions and qualities across 59 cultures: an exploratory analysis

### 2019-11-27 leadership_across_cultures_contexts.Rmd
Adding functions/qual and costs/benefits vars table file.
Adding text in results section.

Across discplines many theoretical models and empirical findings have implicated high intelligence as an important quality positively associated with effective leadership [@johnson_leadership_2015; @judge_intelligence_2004; @neel_being_1980; @antonakis_does_2009]. Recent evidence from within managerial studies suggests a curvilinear effect in that exceptionally intelligence individuals at high end of the distribution may not be the most effective leaders in contemporary business settings [@antonakis_can_2017].

@@

@von_rueden_leadership_2015 outilne four possible mechanisms by which leaders in small-scale societies might offset the potential costs associated with various leadership roles; leaders may, (1) facilitate collective actions that yield net benefits to themselves or their close kin, (2) claim a fee for their services, such a tax or greater share of returns, (3) recieve reciprocal exchange in other currencies from followers who have benefited from their services, or (4) may gain other social or reproductive benefits by signaling their high quality as a mate and social partner.

@@

@@

@@

@@

A logistic PCA on the `r length(quality_vars)` coded leader quality variables, with rank $k=`r dim(logpca_model_qualities$U)[2]`$ optimized by cross-validation, explained `r signif(logpca_model_qualities$prop_deviance_expl, 3)*100`% of the deviance. We draw on the first of the three leader quality components for interpretation and additional analyses and identify this component as a *Prestige vs. Dominance* measure. Low values on Leader qualities PC 1 (Prestige vs. Dominance) are strongly associated with the variables *Experienced/accomplished* and *Knowledgeable/intelligence* and moderately associated with *Oratory skill* and *Innovative*. High values are strongly associated with *Polygnous* and *Coercive authority* and moderately associated with *Many children* and *High quality spouse*. Therefore, the "dominance" side of the *Prestige vs. Dominance* component is also closely liked to reproduction. Measures of physical formidability and aggressiveness are also moderately associated with the "dominance" side of this component. See Figure \@ref(fig:qualitiesPCplot).

@@

Similarly, A logistic PCA on the `r length(quality_vars)` coded leader quality variables, with rank $k=`r dim(logpca_model_functions$U)[2]`$ optimized by cross-validation, `r signif(logpca_model_functions$prop_deviance_expl, 3)*100`% of the deviance. The first Leader functions component distinguishes between interpersonal mediation functions and group organizational and social functions, we define as a *Mediation vs. Organization* measures. Low values on Leader functions PC 1 (Mediation vs. Organization) are strongly associated with the variables *Resolve conflict* and *Provide counsel/direction* and moderately associated with *Political appointments* and *Punishment*, whereas high values are strongly associated with *Social functions* and *Organize cooperation* and moderately associated with *Ritual functions*. See Figure \@ref(fig:functionsPCplot).

@@

Leaders are most commonly receiving benefits in the form of greater access to resources and increased social status. Receiving social services and mating benefits are also relatively frequent benefits of leadership cross-culturally. The costs follow the potential benefits and leaders stand to lose their high social status and greater access to resources. Followers on the other hand most commonly benefit from leadership systems through receiving resources and social services. Mating and fitness benefits are relatively rare for followers. Followers most frequently incur costs in the form of resources lost and in providing social services. For both leaders and followers, individuals stand to lose what they are able to gain. The clear differences lie in leaders' opportunities for increased social status and matting opportunities. See Figure \@ref(fig:varsupportcb).

@@

>>>>

Across disciplines many theoretical models and empirical findings have implicated high intelligence as an important quality positively associated with effective leadership [@johnson_leadership_2015; @judge_intelligence_2004; @neel_being_1980; @antonakis_does_2009]. Recent evidence from within managerial studies suggests a curvilinear effect in that exceptionally intelligence individuals at high end of the distribution may not be the most effective leaders in contemporary business settings [@antonakis_can_2017].

@@

@von_rueden_leadership_2015 outline four possible mechanisms by which leaders in small-scale societies might offset the potential costs associated with various leadership roles; leaders may, (1) facilitate collective actions that yield net benefits to themselves or their close kin, (2) claim a fee for their services, such a tax or greater share of returns, (3) receive reciprocal exchange in other currencies from followers who have benefited from their services, or (4) may gain other social or reproductive benefits by signaling their high quality as a mate and social partner.

@@

To idenfity natural groupings of variables among the research coded leader quality and leader function measures, we conducted two cluster anlayses using the pvclust package [@suzuki_pvclust:_2015]. These analyses allow us to interpret *post-hoc* (1) the degree to which coded variables relate to established theoretical models, (2) identify potential phenotypic profiles of leaders, and (3) offer empirical and theoretical future directions related to suites of qualities or functions likely to covary.

All analyses were conducted with R version...

@@

Within each cluster analyses, we computed the distance between two variables as $1 - cor (v_i,. v_j)$, so that the distance between highly correlated variables is close to $0$, the distance between uncorrelated variables is close to $1$, and the distance between uncorrelated variables us close to $2$. Using the pvclust package [@suzuki_pvclust:_2015] we clustered the variables using the matrix of all pairwise distances using the Ward aggolomeration algorithm.

Figure \@ref(fig:qualcluster) displays a dendrogram from the cluster analysis of the `r length(quality_vars)` ledaer quality variables and Figure \@ref(fig:funcluster) displays a dendrogram from the cluster analysis of the `r length(function_vars)`. Each cluster analysis includes two estimates of significance for how strongly each cluster within the dendrogram is supported by the data. We rely on the AU (Approximately Unbiased) *p* values, which are computed by multiscale bootstrap resampling and represented as percentages (e.g., clusters with AU values $>95$ are strongly supported, and the top-level clusters are automatically highlighted by rectangles).

@@

The leader quality cluster analysis (Figure \@ref(fig:qualcluster)) idenfitified two top-level clusters. Within each top-level cluster we identify two interpretable and generally strongly supported sub-clusters which are annotated with text. Within the first top-level cluster, we identify a dominant/shaman cluster, which suggests a strong link between dominance-based leadership (e.g., *Authority*, *Aggressiveness*, *Feared*) and shamanism (e.g., *Supernatural*, *Artistic performance*). The second sub-cluster includes qualities of leaders related to material, social, and reproductive success.

@@

A logistic PCA on the `r length(quality_vars)` coded leader quality variables, with rank $k = `r dim(logpca_model_qualities$U)[2]`$ and  $m = `r logpca_model_qualities$m`$ optimized by cross-validation, explained `r signif(logpca_model_qualities$prop_deviance_expl, 3)*100`% of the deviance. We draw on the first two leader quality components for interpretation and additional analyses. We identify the first component (PC1) as a *Prestige vs. Dominance* measure. Low values on Leader qualities PC 1 are strongly associated with the variables *Experienced/accomplished* and *Knowledgeable/intelligence* and moderately associated with *Confidence* and *Bravery*. High values are strongly associated with *Polygnous* and *Authority* and moderately associated with *High quality spouse*. Therefore, the "dominance" side of the *Prestige vs. Dominance* component is also closely liked to reproduction, as demonstrated in the cluster analysis (Figure \@ref(fig:qualcluster)). Measures of physical formidability and aggressiveness are also moderately associated with the "dominance" side of this component. See Figure \@ref(fig:qualitiesPCplot). The second leader qualities component we identify as an *In-group favoritism vs. Social Status* measure. Low values on the Leader qualities PC 2 are strongly associated with *Aggressiveness*, *Bravery*, *Xenophobia*, and *Loyalty*. High Values are strongly associated with *High status* and moderately associated with *Fairness* and *Age* and *Decisiveness/decision-making*.

@@

Similarly, A logistic PCA on the `r length(function_vars)` coded leader function variables, with rank $k=`r dim(logpca_model_functions$U)[2]`$ and $m = `r logpca_model_functions$m`$ optimized by cross-validation, `r signif(logpca_model_functions$prop_deviance_expl, 3)*100`% of the deviance. The first Leader functions component distinguishes between social functions and general prosocial function from other more specific leader functions, which we identify as a *Social functions* component. High values on Leader function PC 1 are strongly associated with *Social functions*. The Leader functions PC 2 distinguished between organizational and mediation-based functions, which we identify as a *Organization vs. Mediation* component. Low values on Leader functions PC 2 are strongly associated with the variables *Organize cooperation* and *Political appointments*, whereas high values are strongly associated with *Resolve conflict* and moderately associated with *Provide counsel/direction*. See Figure \@ref(fig:functionsPCplot).

@@

Leaders are most commonly receiving benefits in the form of greater access to mateiral resources and increased social status. Receiving social services, mating benefits, and food benefits are also relatively frequent benefits of leadership cross-culturally. The costs follow the potential benefits and leaders stand to lose their high social status and greater access to resources.

Followers on the other hand also benefit from leadership systems through receiving material resources but otherwise most frequently benefit through social *services* provided. Mating and fitness benefits are relatively rare for followers. Followers most frequently incur costs in the form of material resources lost and in providing social services. For both leaders and followers, individuals stand to lose what they are able to gain. The clear differences lie in leaders' opportunities for increased social status and matting opportunities. See Figure \@ref(fig:varsupportcb).

@@

## Limitations
